---
title: "HW 11"
author: "Yujui Chang"
date: "`r Sys.Date()`"
output: html_document
urlcolor: "blue"
params:
  solutions: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = params$solutions, 
                      eval       = params$solutions,
                      fig.align  = "center",
                      fig.height = 3, 
                      fig.width  = 5)
```

# Instructions

- Write your solutions in this starter file. You should modify the 
  "author" field in the YAML header.
- Do not modify the paths of any files.
- Only commit R Markdown and HTML files (no PDF or Word files). Make sure
  you have knitted to HTML for your final submission.
- **Make sure to commit each time you answer a question.** Lack of informative
  and frequent commits will result in point deductions.
- Only include the necessary code, not any extraneous code, to answer the 
  questions.
- For this homework, **use only data.table and not the tidyverse**.
- Learning objectives:
    - Manipulating data.tables.
    - Tidying data.tables.

# Bob Ross

The data in "bob.csv" contains information on episodes hosted
by [Bob Ross](https://en.wikipedia.org/wiki/Bob_Ross), a painter who hosted a 
popular television series on PBS. Each episode 
would consist of him completing an entire painting. His paintings almost 
always consisted heavily of elements from nature: trees, clouds, mountains, 
lakes, etc. These data consist of indicators for what elements are in each 
episode. This dataset was taken from the excellent crew at
[fivethirtyeight](https://github.com/fivethirtyeight/data/tree/master/bob-ross). 
See [here](https://fivethirtyeight.com/features/a-statistical-analysis-of-the-work-of-bob-ross/) 
for their article. The variables are:

- `EPISODE`: The season and episode number of the episode.
- `TITLE`: The title of the painting.
- Every other variable is an indicator for whether the episode contains 
  the element described by the variable name. For example, `BARN` 
  is `0` if the episode does not have a barn in the painting and `1` if the 
  episode does have a barn in the painting.

1. Use data.table and `fread()` to read these data into R.
```{r}
library(data.table)
library(readr)
library(tidyverse)
bob <- fread("../data/bob.csv")
str(bob)
```
    
2. Tidy these data by having new columns called "`ELEMENT`" and "`PRESENT`". 
   `ELEMENT` should contain the name of the element included in the painting
   and the variable `PRESENT` should be `1` if that element is present
   and `0` otherwise. Your new `data.table` should look like this:
```{r}
bob <- melt(data = bob, 
            id.vars = c("EPISODE", "TITLE"), 
            measure.vars = colnames(bob)[3:69], 
            variable.name = "ELEMENT", 
            value.name = "PRESENT")
bob
```
    
3. The variable `EPISODE` contains both the season number and the episode number. 
   Create two new variables called "`SEASON`" and "`EPISODE_NUM`". `SEASON`
   should be a numeric containing the season number and `EPISODE_NUM` should
   be a numeric containing the episode number within the season. 
   Do this by reference. You can use readr parsers here.
   Your data.table should look like this:
```{r}
bob[,c("SEASON", "EPISODE_NUM"):= tstrsplit(EPISODE, split="E")]
bob[,SEASON := parse_number(SEASON)]
bob[,EPISODE_NUM := parse_number(EPISODE_NUM)]

head(bob)
tail(bob)
```
    
4. Recode the `PRESENT` variable to have values that are more human-readable.
   Do this by reference.
```{r}
bob[, PRESENT:= as.character(PRESENT)]

bob[PRESENT=="0", PRESENT := "UNPRESENT"]
bob[PRESENT=="1", PRESENT := "PRESENT"]

bob
```

5. Calculate the proportion of episodes each season that have each element. Save
   these summaries in a new data.table. Arrange the rows in decreasing
   order of proportion. Your data.table should look like this:
```{r}
newbob <- bob
newbob[PRESENT=="UNPRESENT", PRESENT := "0"]
newbob[PRESENT=="PRESENT", PRESENT := "1"]
newbob[, PRESENT:= parse_number(PRESENT)]
newbob


newbob[, .(PROPORTION = mean(PRESENT)), by= .(ELEMENT,SEASON)][order(-PROPORTION)]

```

6. Did Bob draw trees less frequently over the seasons. Explore with an 
   appropriate plot. You can use ggplot2 or base R.
```{r}
bob_tree <- bob[ELEMENT=="TREE", .(drawtree = sum(PRESENT)), by=.(SEASON)]
bob_tree[, SEASON:= as.factor(SEASON)]

bob_tree %>%
   ggplot(aes(x= SEASON, y= drawtree))+
   geom_point()+
   geom_smooth(method="lm", se= FALSE)+
   theme_bw()

```

7. Does Bob prefer conifers or deciduous trees? Justify your conclusions with
   numerical summaries and a statistical test. 
```{r}
bob[ELEMENT == "CONIFER" | ELEMENT == "DECIDUOUS", .(MEAN_TREE = mean(PRESENT)), by= .(ELEMENT)]
# Both two trees have similar mean. It is hard to tell Bob prefer which trees.

bob_con <- bob[ELEMENT == "CONIFER" | ELEMENT == "DECIDUOUS", .(ELEMENT, PRESENT)]

t.test(PRESENT~ ELEMENT, bob_con)

# P-value is big (0.2893), we have very weak evidence to say that Bob prefer which trees.

```

